Comb associative operads

### 2018-06-05 Article.tex
Petites corrections

% Author: Cyrille Chenavier, Christophe Cordero, Samuele Giraudo
% Creation: may. 2018
% Modifications: may. 2018

@@

\newcommand{\Zero}{\mathtt{0}}
\newcommand{\Two}{\mathtt{2}}
\newcommand{\HilbertSeries}{\mathcal{H}}
\newcommand{\GeneratingSet}{\mathfrak{G}}
\newcommand{\FreeOperad}{\mathrm{F}}
\newcommand{\Mag}{\mathsf{Mag}}
\newcommand{\As}{\mathsf{As}}
\newcommand{\CAs}[1]{\mathsf{CAs}^{(#1)}}
\newcommand{\PrefixWord}{\mathrm{p}}
\newcommand{\Deg}{\mathrm{deg}}
\newcommand{\LC}[1]{\text{LC}^{(#1)}}
\newcommand{\RC}[1]{\text{RC}^{(#1)}}

\DeclareMathOperator{\Product}{\star}
\DeclareMathOperator{\Congr}{\equiv}
\DeclareMathOperator{\Rew}{\to}
\DeclareMathOperator{\RewRT}{\overset{*}{\Rew}}
\DeclareMathOperator{\RewT}{\overset{+}{\Rew}}
\DeclareMathOperator{\RewContext}{\Rightarrow}
\DeclareMathOperator{\RewContextRT}{\overset{*}{\RewContext}}
\DeclareMathOperator{\RewContextT}{\overset{+}{\RewContext}}

@@

Let us now define for any $d \geq 1$ the \Def{$\gamma$-comb associative
operad} $\CAs{\gamma}$ as the quotient operad $\Mag/_{\CongrCAs{\gamma}}$ where
$\CongrCAs{\gamma}$ is the smallest operad congruence of $\Mag$
satisfying
\begin{equation} \label{equ:congruence_Asd}
    \underbrace{(\dots (\Product \circ_1 \Product) \circ_1 \dots)
        \circ_1 \Product}
    _{\gamma \mbox{ \footnotesize operands}}
    \enspace \CongrCAs{\gamma} \enspace
    \underbrace{\Product \circ_2
        (\dots \circ_2 (\Product \circ_2 \Product) \dots)}
    _{\gamma \mbox{ \footnotesize operands}}.
\end{equation}
In words, \eqref{equ:congruence_Asd} says that the left and the right
comb binary trees of degree $\gamma$ are equivalent for $\CongrCAs{\gamma}$.
Notice that $\CongrCAs{1}$ is trivial so that $\CAs{1} = \Mag$ and that
$\CongrCAs{2}$ is the operad congruence defined
by~\eqref{equ:congruence_as} so that $\CAs{2} = \As$.

@@

We denote by CAs the set of $\gamma$-comb associative operads where $\gamma$ belongs
the set of strictly positive integers:
\[\text{CAs}=\left\{\CAs{\gamma}\mid\gamma\geq 1\right\}.\]
Our purpose is to show that CAs admits a lattice structure. For that, we provide
a description of operads morphisms between the elements of CAs.

\begin{Proposition}\label{prop:morphisms_CAs_d}

  Let $\gamma$ and $\gamma'$ be two strictly positive integers. There exists at most
  one operads morphism $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$. This morphism
  exists if and only if $\emph{LC}\left(\gamma'\right)\CongrCAs{\gamma}\emph{RC}\left(\gamma'\right)$ and it is surjective in this
  case. Moreover, $\varphi$ is injective if and only if $\gamma=\gamma'$, that is if and only
  if $\varphi$ is the identity morphism.

  \end{Proposition}

\begin{proof}
  The operad $\CAs{\gamma'}$ is generated by one binary generator $t^{(\gamma')}$, so that
  $\varphi$ is entirely determined by $\varphi\left(t^{(\gamma')}\right)$. Moreover,
  $\varphi\left(t^{(\gamma')}\right)$ has to be of arity $2$ in $\CAs{\gamma}$, so that
  $\varphi\left(t^{(\gamma')}\right)$ has to be equal to $t^{(\gamma)}$. Hence, if $\varphi$
  exists, it is the unique operads morphism from $\CAs{\gamma'}$ and $\CAs{\gamma}$,
  and in this case, $t^{(\gamma)}$ being in the image of $\varphi$, the latter is
  surjective. Moreover, $\varphi$ is well defined if and only if
  $\varphi\left(\LC{\gamma'}\right)$ and $\varphi\left(\RC{\gamma'}\right)$ are equal in
  $\CAs{\gamma}$, that is if and only if $\LC{\gamma'}\CongrCAs{\gamma}\RC{\gamma'}$. In particular,
  if $\gamma'$ is strictly smaller than $\gamma$, then there does not exist any morphism
  $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$. Moreover, if $\gamma'$ is strictly smaller
  than $\gamma$, the cardinality of $\CAs{\gamma'}\left(\gamma\right)$ is equal to the cardinality
  of $\CAs{\gamma}(\gamma)$ plus one. Hence, if there exists an injective morphism
  $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$, we must have $\gamma=\gamma'$, and in this case,
  $\varphi$ is the identity morphism.
\end{proof}

We define the binary relation $\preceq$ on CAs as follows: we have $\CAs{\gamma}\preceq\CAs{\gamma'}$
if and only if there exists a morphism $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$.

\begin{Proposition}\label{prop:poset_CAs}
  The binary relation $\preceq$ is a partially order on \emph{CAs}, that is
  $\left(\emph{CAs},\ \preceq\right)$ is a poset.
  \end{Proposition}

\begin{proof}
  The binary relation $\preceq$ is reflexive since there exists the identity morphism
  of $\CAs{\gamma}$ for every strictly positive integer $\gamma$. It is transitive since if
  there exist two morphisms  $\varphi:\CAs{\gamma}\to\CAs{\gamma'}$ and
  $\psi:\CAs{\gamma''}\to\CAs{\gamma'}$, then $\psi\circ\varphi$ is a morphism from
  $\CAs{\gamma''}$ to $\CAs{\gamma}$. Now, let us assume that there exist morphisms
  $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$ and $\psi:\CAs{\gamma}\to\CAs{\gamma'}$.
  In particular, $\psi\circ\varphi$ and $\varphi\circ\psi$ are endomorphisms of $\CAs{\gamma'}$
  and $\CAs{\gamma}$, respectively. From Proposition~\ref{prop:morphisms_CAs_d}, these two
  morphisms are identity morphisms, so that $\varphi$ and $\psi$ are injective. From
  Proposition~\ref{prop:morphisms_CAs_d}, $\gamma$ and $\gamma'$ are equal, which proves that
  $\preceq$ is anti-symmetric, hence a partial order.
  \end{proof}

In order to show that the poset $\left(\text{CAs},\ \preceq\right)$ extends into a lattice,
that is two elements of CAs admit lower-bounds and upper-bounds, we relate
$\left(\text{CAs},\ \preceq\right)$ to the lattice of integers
$\left(\N,\ \mid,\ \text{gcd},\ \text{lcm}\right)$, where $\mid$ denotes the division
relation, gcd denotes the \emph{greatest common divisor} and lcm the \emph{least common multiple}
operators, respectively. In order to obtain this relationship, we introduce the notion of
\Def{left rank} of an element $t\in\Mag$: this is the length $\text{lr}\left(t\right)$ of
the left branch beginning at the root of $t$. The left rank of $t$ is represented as follows:
\[\textbf{PICTURE}\]

>>>>

% Author: Cyrille Chenavier, Christophe Cordero, Samuele Giraudo
% Creation: may. 2018
% Modifications: may. 2018, june 2018

@@

\newcommand{\Zero}{\mathtt{0}}
\newcommand{\Two}{\mathtt{2}}
\newcommand{\HilbertSeries}{\mathcal{H}}
\newcommand{\GeneratingSet}{\mathfrak{G}}
\newcommand{\FreeOperad}{\mathrm{F}}
\newcommand{\Mag}{\mathsf{Mag}}
\newcommand{\As}{\mathsf{As}}
\newcommand{\CAs}[1]{\mathsf{CAs}^{(#1)}}
\newcommand{\CAsAll}{\mathsf{CAs}}
\newcommand{\PrefixWord}{\mathrm{p}}
\newcommand{\Deg}{\mathrm{deg}}
\newcommand{\LC}[1]{\mathfrak{lc}^{(#1)}}
\newcommand{\RC}[1]{\mathfrak{lr}^{(#1)}}
\newcommand{\LBranch}{\mathrm{lb}}
\newcommand{\RBranch}{\mathrm{rb}}

\DeclareMathOperator{\Product}{\star}
\DeclareMathOperator{\Congr}{\equiv}
\DeclareMathOperator{\Rew}{\to}
\DeclareMathOperator{\RewRT}{\overset{*}{\Rew}}
\DeclareMathOperator{\RewT}{\overset{+}{\Rew}}
\DeclareMathOperator{\RewContext}{\Rightarrow}
\DeclareMathOperator{\RewContextRT}{\overset{*}{\RewContext}}
\DeclareMathOperator{\RewContextT}{\overset{+}{\RewContext}}
\DeclareMathOperator{\lcm}{lcm}

@@

Let us now define for any $\gamma \geq 1$ the \Def{$\gamma$-comb
associative operad} $\CAs{\gamma}$ as the quotient operad
$\Mag/_{\CongrCAs{\gamma}}$ where $\CongrCAs{\gamma}$ is the smallest
operad congruence of $\Mag$ satisfying
\begin{equation} \label{equ:congruence_Asd}
    \underbrace{(\dots (\Product \circ_1 \Product) \circ_1 \dots)
        \circ_1 \Product}
    _{\gamma \mbox{ \footnotesize operands}}
    \enspace \CongrCAs{\gamma} \enspace
    \underbrace{\Product \circ_2
        (\dots \circ_2 (\Product \circ_2 \Product) \dots)}
    _{\gamma \mbox{ \footnotesize operands}}.
\end{equation}
In words, \eqref{equ:congruence_Asd} says that the left and the right
comb binary trees of degree $\gamma$ are equivalent for
$\CongrCAs{\gamma}$. Notice that $\CongrCAs{1}$ is trivial so that
$\CAs{1} = \Mag$ and that $\CongrCAs{2}$ is the operad congruence
defined by~\eqref{equ:congruence_as} so that $\CAs{2} = \As$.

@@

Let
\begin{equation}
    \CAsAll := \left\{\CAs{\gamma} : \gamma\geq 1\right\}
\end{equation}
be the set of all the $\gamma$-comb associative operads
Our purpose is to show that $\CAsAll$ admits a lattice structure. For
that, we provide a description of operads morphisms between the
elements of $\CAsAll$.

\begin{Proposition}\label{prop:morphisms_CAs_d}
    Let $\gamma$ and $\gamma'$ be two positive integers. There exists at
    most one operad morphism $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$.
    This  morphism exists if and only if
    \begin{math}
        \LC{\gamma'}\CongrCAs{\gamma}\RC{\gamma'}
    \end{math}
    and it is surjective in this case. Moreover, $\varphi$ is injective
    if and only if $\gamma = \gamma'$, that is if and only if
    $\varphi$ is the identity morphism.
\end{Proposition}
\begin{proof}
    The operad $\CAs{\gamma'}$ is generated by one binary generator
    $\Product$, so that $\varphi$ is entirely determined by
    $\varphi\left(\Product\right)$. Moreover, $\varphi(\Product)$ has
    to be of arity $2$ in $\CAs{\gamma}$, so that $\varphi(\Product)$
    has to be equal to $\Product \in \CAs{\gamma}$. Hence, if $\varphi$
    exists, it is the unique operad morphism from $\CAs{\gamma'}$ and
    $\CAs{\gamma}$, and in this case, $\Product \in \CAs{\gamma'}$
    being in the image of $\varphi$, the latter is surjective.
    Moreover, $\varphi$ is well defined if and only if
    $\varphi\left(\LC{\gamma'}\right)$ and
    $\varphi\left(\RC{\gamma'}\right)$ are equal in $\CAs{\gamma}$,
    that is if and only if
    $\LC{\gamma'} \CongrCAs{\gamma} \RC{\gamma'}$. In particular, if
    $\gamma'$ is smaller than $\gamma$, then there does not exist any
    morphism $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$.
    \Todo{Faux car on peut avoir $\gamma' = 1$, puisque
    $\LC{1} = \RC{1}$.}
    Moreover, if $\gamma'$ is smaller than $\gamma$, the cardinality of
    $\CAs{\gamma'}\left(\gamma\right)$ is equal to the cardinality of
    $\CAs{\gamma}(\gamma)$ plus one. Hence, if there exists an injective
    morphism $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$, we must have
    $\gamma = \gamma'$, and in this case, $\varphi$ is the identity
    morphism.
\end{proof}

We define the binary relation $\preceq$ on $\CAsAll$ as follows: we have
$\CAs{\gamma}\preceq\CAs{\gamma'}$ if and only if there exists a
morphism $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$.

\begin{Proposition}\label{prop:poset_CAs}
    The binary relation $\preceq$ is a partially order on $\CAsAll$,
    that is $\left(\CAsAll,\ \preceq\right)$ is a poset.
\end{Proposition}
\begin{proof}
    The binary relation $\preceq$ is reflexive since there exists the
    identity morphism of $\CAs{\gamma}$ for every positive integer
    $\gamma$. It is transitive since if there exist two morphisms
    $\varphi:\CAs{\gamma}\to\CAs{\gamma'}$ and
    $\psi:\CAs{\gamma''}\to\CAs{\gamma'}$, then $\psi\circ\varphi$ is a
    morphism from $\CAs{\gamma''}$ to $\CAs{\gamma}$. Now, let us assume
    that there exist morphisms $\varphi:\CAs{\gamma'}\to\CAs{\gamma}$
    and $\psi:\CAs{\gamma}\to\CAs{\gamma'}$. In particular,
    $\psi\circ\varphi$ and $\varphi\circ\psi$ are endomorphisms of
    $\CAs{\gamma'}$ and $\CAs{\gamma}$, respectively. From
    Proposition~\ref{prop:morphisms_CAs_d}, these two morphisms are
    identity morphisms, so that $\varphi$ and $\psi$ are injective. From
    Proposition~\ref{prop:morphisms_CAs_d}, $\gamma$ and $\gamma'$ are
    equal, which proves that $\preceq$ is anti-symmetric, hence a
    partial order.
\end{proof}

In order to show that the poset $\left(\CAsAll,\ \preceq\right)$
extends into a lattice, that is two elements of $\CAsAll$ admit
lower-bounds and upper-bounds, we relate
$\left(\CAsAll,\ \preceq\right)$ to the lattice of integers
$\left(\N,\ \mid,\ \gcd,\ \lcm\right)$, where $\mid$
denotes the division relation, $\gcd$ denotes the \emph{greatest common divisor} and $\lcm$ the \emph{least common multiple}
operators, respectively. In order to obtain this relationship, we introduce the notion of
\Def{left rank} of an element $\Tfr \in\Mag$: this is the number
$\LBranch(\Tfr)$ of internal nodes in the left branch
beginning at the root of $\Tfr$. The left rank of $\Tfr$ is represented as follows:
\[\textbf{PICTURE}\]
